---
title: "Spectometer Regression" 
author: "**Siyu Jian / sj9va**"
date: " Wed Mar 10 | 10:55am"
output: R6018::homework
---

*******************************************
```{r config, echo=FALSE}
source(system.file("config/hw_config.R", package="R6018")) # knitr settings
# options(dplyr.summarise.inform = FALSE)  # ignore dplyr message about grouping
```


# Required R packages and Directories

### {.solution}
```{r packages, message=FALSE, warning=FALSE}
data.dir = 'https://mdporter.github.io/SYS6018/data/' # data directory
library(R6018)     # functions for SYS-6018
library(tidyverse) # functions for data manipulation   
library(mlbench)
library(glmnet)
library(glmnetUtils) 
```

# Load the focal plane training dataset 
```{r warning=FALSE}
# focal_url = "http://localhost/Data/spectro_nn/focalPlane/Equal/EqEvt731/order7_ep5/combine.csv"
focal_url = "http://localhost/Data/spectro_nn/focalPlane/Equal/EqEvt731/order10_ep5/combine.csv"

dataset_focal = readr::read_csv(focal_url)
dataSet = dataset_focal %>% as.matrix()
# dataSet
```

# Spectrometer Regression {.tabset} 
## 20-fold lasso regression 
### 1. training the model
```{r }

trainSize = round(0.7*nrow(dataSet)) 

train  = sample(nrow(dataSet),size = trainSize)
test   = - train

trainSet = dataset_focal[train,]
train_X = trainSet %>% select(-evtID, -runID, -CutID,-SieveRowID,-SieveColID,-bpmX,-bpmY,-targCalTh,-targCalPh)
train_y_theta = trainSet %>% select(targCalTh)
train_y_phi = trainSet %>% select(targCalPh)


n.fold = 20
fold = sample(rep(1:n.fold, length=nrow(train_X)))
fit.theta = cv.glmnet(as.matrix(train_X),as.matrix(train_y_theta),foldid=fold)
fit.phi   = cv.glmnet(as.matrix(train_X),as.matrix(train_y_phi),foldid = fold)
```

```{r}

plot(fit.phi,log='y')
plot(fit.theta,log='y')
```
### 2. check how it behave on the training dataset 
```{r}

trainSetRes  = trainSet %>% select(evtID, runID, CutID,SieveRowID,SieveColID,bpmX,bpmY,targCalTh,targCalPh)

train_ResX   = trainSet %>% select(-evtID, -runID, -CutID,-SieveRowID,-SieveColID,-bpmX,-bpmY,-targCalTh,-targCalPh)
train_ResX_theta = trainSet  %>% select(targCalTh)
train_ResX_phi   = trainSet  %>% select(targCalPh)

train.y.theta.pred = predict(fit.theta,newx = as.matrix(train_ResX),s="lambda.1se")
colnames(train.y.theta.pred) = c("pred.theta")
trainSetRes <- cbind(trainSetRes,pred.thet = train.y.theta.pred)

train.y.phi.pred = predict(fit.phi,newx = as.matrix(train_ResX),s="lambda.1se")
colnames(train.y.phi.pred) = c("pred.phi")
trainSetRes <- cbind(trainSetRes,pred.phi = train.y.phi.pred)
trainSetRes$resid.theta <- trainSetRes$targCalTh - trainSetRes$pred.theta 
trainSetRes$resid.phi <- trainSetRes$targCalPh - trainSetRes$pred.phi 


```

```{r}
val = trainSetRes %>%
  filter(.$runID == 2239)
ggplot(val) +geom_bin2d(aes(x=pred.phi,y=pred.theta,color = "blue"),bins = 300) + geom_point(aes(y=targCalTh,x=targCalPh,color="green")) + 
  xlim(-0.015,0.025) + 
  labs(title=sprintf("run_2d_%d",2239))


val = trainSetRes %>%
  filter(.$runID == 2240)

ggplot(val) +geom_bin2d(aes(x=pred.phi,y=pred.theta,color = "blue"),bins = 300) + geom_point(aes(y=targCalTh,x=targCalPh,color="green")) + 
  xlim(-0.015,0.025) + 
  labs(title=sprintf("run_2d_%d",2240))


val = trainSetRes %>%
  filter(.$runID == 2241)
ggplot(val) +geom_bin2d(aes(x=pred.phi,y=pred.theta,color = "blue"),bins = 300) + geom_point(aes(y=targCalTh,x=targCalPh,color="green")) + 
  xlim(-0.015,0.025) + 
  labs(title=sprintf("run_2d_%d",2241))

val = trainSetRes %>%
  filter(.$runID == 2244)
ggplot(val) +geom_bin2d(aes(x=pred.phi,y=pred.theta,color = "blue"),bins = 300) + geom_point(aes(y=targCalTh,x=targCalPh,color="green")) + 
  xlim(-0.015,0.025) + 
  labs(title=sprintf("run_2d_%d",2244))

val = trainSetRes %>%
  filter(.$runID == 2245)
ggplot(val) +geom_bin2d(aes(x=pred.phi,y=pred.theta,color = "blue"),bins = 300) + geom_point(aes(y=targCalTh,x=targCalPh,color="green")) + 
  xlim(-0.015,0.025) + 
  labs(title=sprintf("run_2d_%d",2245))


val = trainSetRes %>%
  filter(.$runID == 2256)

ggplot(val) +geom_bin2d(aes(x=pred.phi,y=pred.theta,color = "blue"),bins = 300) + geom_point(aes(y=targCalTh,x=targCalPh,color="green")) + 
  xlim(-0.015,0.025) + 
  labs(title=sprintf("run_2d_%d",2256))


val = trainSetRes %>%
  filter(.$runID == 2257)

ggplot(val) +geom_bin2d(aes(x=pred.phi,y=pred.theta,color = "blue"),bins = 300) + geom_point(aes(y=targCalTh,x=targCalPh,color="green")) + 
  xlim(-0.015,0.025) + 
  labs(title=sprintf("run_2d_%d",2257))
```

### 3.plot the validation plot 
```{r}
# make prediction on the theta and phi angles
testSet = dataset_focal[test,]

testSetRes = testSet %>% select(evtID, runID, CutID,SieveRowID,SieveColID,bpmX,bpmY,targCalTh,targCalPh)

test_X = testSet %>% select(-evtID, -runID, -CutID,-SieveRowID,-SieveColID,-bpmX,-bpmY,-targCalTh,-targCalPh)
test_y_theta = testSet %>% select(targCalTh)
test_y_phi   = testSet %>% select(targCalPh)

test.y.theta.pred = predict(fit.theta,newx = as.matrix(test_X),s="lambda.1se")
colnames(test.y.theta.pred) = c("pred.theta")
testSetRes <- cbind(testSetRes,pred.thet = test.y.theta.pred)

test.y.phi.pred = predict(fit.phi,newx = as.matrix(test_X),s="lambda.1se")
colnames(test.y.phi.pred) = c("pred.phi")
testSetRes <- cbind(testSetRes,pred.phi = test.y.phi.pred)
testSetRes$resid.theta <- testSetRes$targCalTh - testSetRes$pred.theta 
testSetRes$resid.phi <- testSetRes$targCalPh - testSetRes$pred.phi 

# testSetRes
```




```{r}
unique(testSetRes$runID) %>% print()
val = testSetRes %>% filter(.$runID == 2256)
val 
```

```{r}
unique(testSetRes$runID)

val = testSetRes %>%
  filter(.$runID == 2239)
ggplot(val) +geom_bin2d(aes(x=pred.phi,y=pred.theta,color = "blue"),bins = 500) + geom_point(aes(y=targCalTh,x=targCalPh,color="green")) + 
  xlim(-0.015,0.025) + 
  labs(title=sprintf("run_2d_%d",2239))


val = testSetRes %>%
  filter(.$runID == 2240)

ggplot(val) +geom_bin2d(aes(x=pred.phi,y=pred.theta,color = "blue"),bins = 500) + geom_point(aes(y=targCalTh,x=targCalPh,color="green")) + 
  xlim(-0.015,0.025) + 
  labs(title=sprintf("run_2d_%d",2240))


val = testSetRes %>%
  filter(.$runID == 2241)
ggplot(val) +geom_bin2d(aes(x=pred.phi,y=pred.theta,color = "blue"),bins = 500) + geom_point(aes(y=targCalTh,x=targCalPh,color="green")) + 
  xlim(-0.015,0.025) + 
  labs(title=sprintf("run_2d_%d",2241))

val = testSetRes %>%
  filter(.$runID == 2244)
ggplot(val) +geom_bin2d(aes(x=pred.phi,y=pred.theta,color = "blue"),bins = 500) + geom_point(aes(y=targCalTh,x=targCalPh,color="green")) + 
  xlim(-0.015,0.025) + 
  labs(title=sprintf("run_2d_%d",2244))

val = testSetRes %>%
  filter(.$runID == 2245)
ggplot(val) +geom_bin2d(aes(x=pred.phi,y=pred.theta,color = "blue"),bins = 500) + geom_point(aes(y=targCalTh,x=targCalPh,color="green")) + 
  xlim(-0.015,0.025) + 
  labs(title=sprintf("run_2d_%d",2245))


val = testSetRes %>%
  filter(.$runID == 2256)

ggplot(val) +geom_bin2d(aes(x=pred.phi,y=pred.theta,color = "blue"),bins = 500) + geom_point(aes(y=targCalTh,x=targCalPh,color="green")) + 
  xlim(-0.015,0.025) + 
  labs(title=sprintf("run_2d_%d",2256))


val = testSetRes %>%
  filter(.$runID == 2257)

ggplot(val) +geom_bin2d(aes(x=pred.phi,y=pred.theta,color = "blue"),bins = 500) + geom_point(aes(y=targCalTh,x=targCalPh,color="green")) + 
  xlim(-0.015,0.025) + 
  labs(title=sprintf("run_2d_%d",2257))

#ggplot(val,aes(x=pred.phi,y=pred.theta,color="red")) + geom_point()

```

```{r}
residalRes = tibble(runID = numeric(), cutID = numeric(),mean.resid.theta = numeric(),stde.resid.theta = numeric(),mean.resid.phi = numeric(),stde.resid.phi= numeric())
runID.list = unique(testSetRes$runID)

for (id in runID.list) {
    test.res = testSetRes %>%
      filter(.$runID == id)
    
    cutid.list = unique(test.res$CutID)
    for (cutid in cutid.list){
      test.res.cutid = test.res %>%
        filter(.$CutID == cutid)
      mean.resid.theta = mean(test.res.cutid$resid.theta)
      stde.resid.theta = sd(test.res.cutid$resid.theta)/sqrt(length(test.res.cutid$resid.theta))
      
      mean.resid.phi = mean(test.res.cutid$resid.phi)
      stde.resid.phi = sd(test.res.cutid$resid.phi)/sqrt(length(test.res.cutid$resid.phi))
      residalRes = add_row(residalRes,tibble(runID=id,cutID = cutid, mean.resid.theta = mean.resid.theta,
                                             stde.resid.theta=stde.resid.theta,
                                             mean.resid.phi = mean.resid.phi,
                                             stde.resid.phi = stde.resid.phi))
    }
}



residalRes %>%
  filter(.$runID == 2239) %>%
  ggplot(aes(x=cutID,y=mean.resid.phi)) + geom_point() + ylim(-0.003,0.003) +  
  labs(title=sprintf("residual_%d",2239))


residalRes %>%
  filter(.$runID == 2240) %>%
  ggplot(aes(x=cutID,y=mean.resid.phi)) + geom_point() + ylim(-0.003,0.003) +  
  labs(title=sprintf("residual_%d",2240))

residalRes %>%
  filter(.$runID == 2241) %>%
  ggplot(aes(x=cutID,y=mean.resid.phi)) + geom_point() + ylim(-0.003,0.003) +  
  labs(title=sprintf("residual_%d",2241))

residalRes %>%
  filter(.$runID == 2244) %>%
  ggplot(aes(x=cutID,y=mean.resid.phi)) + geom_point() + ylim(-0.003,0.003) +  
  labs(title=sprintf("residual_%d",2244))


residalRes %>%
  filter(.$runID == 2245) %>%
  ggplot(aes(x=cutID,y=mean.resid.phi)) + geom_point() + ylim(-0.003,0.003) +  
  labs(title=sprintf("residual_%d",2245))

residalRes %>%
  filter(.$runID == 2256) %>%
  ggplot(aes(x=cutID,y=mean.resid.phi)) + geom_point() + ylim(-0.003,0.003) +  
  labs(title=sprintf("residual_%d",2256))

residalRes %>%
  filter(.$runID == 2257) %>%
  ggplot(aes(x=cutID,y=mean.resid.phi)) + geom_point() + ylim(-0.003,0.003) +  
  labs(title=sprintf("residual_%d",2257))

res = testSetRes %>%
  filter(.$runID == 2239)

mid_phi_df <- res %>%
  group_by(CutID) %>%
  summarize(median=median(resid.phi))

ggplot(res, aes(x=resid.phi,color = "CutID")) + geom_density() + 
  geom_vline(data =mid_phi_df, aes(xintercept = median))+ 
  xlim(-0.002,0.002)
  
```



## n.fold cross validation regression

### 1.prepare the ridge regression dataset
```{r cache=TRUE}
# n.folds = seq(10,100,by=30)
# fold.scan.res = tibble(fold = numeric(),type = str_c(),lambda.min = numeric(),lambda.1se = numeric(), mse.min = numeric(),mse.1se = numeric())
# 
# ptm <- proc.time()
# 
# for (n.fold in n.folds){
#   fold = sample(rep(1:n.fold,length = nrow(train_X)))
#   fit.theta.step = cv.glmnet(as.matrix(train_X),as.matrix(train_y_theta),foldid = fold)
#   fit.phi.step   = cv.glmnet(as.matrix(train_X),  as.matrix(train_y_phi),foldid = fold)
# 
#   # get the min mse
#   mse.min.theta = fit.theta.step$cvm[fit.theta.step$lambda == fit.theta.step$lambda.min]
#   mse.1se.theta = fit.theta.step$cvm[fit.theta.step$lambda == fit.theta.step$lambda.1se]
#   lambda.min.theta = fit.theta.step$lambda.min
#   lambda.1se.theta = fit.theta.step$lambda.1se
#   
#   # res.theta = tibble(fold = n.fold, type = "theta",
#   #                     lambda.min = lambda.min.theta,
#   #                     lambda.1se = lambda.1se.theta,
#   #                     mse.min = mse.min.theta,
#   #                     mse.1se = mse.1se.theta )
#   fold.scan.res = add_row(fold.scan.res, fold = n.fold, type = "theta",
#                       lambda.min = lambda.min.theta,
#                       lambda.1se = lambda.1se.theta,
#                       mse.min = mse.min.theta,
#                       mse.1se = mse.1se.theta)
#   
#   # get the phi information
#   mse.min.phi = fit.phi.step$cvm[fit.phi.step$lambda == fit.phi.step$lambda.min]
#   mse.1se.phi = fit.phi.step$cvm[fit.phi.step$lambda == fit.phi.step$lambda.1se]
#   lambda.min.phi = fit.phi.step$lambda.min 
#   lambda.1se.phi = fit.phi.step$lambda.1se
#   
#   # res.phi = tibble(fold = n.fold, type = "phi",
#   #                  lambda.min = lambda.min.phi,
#   #                  lambda.1se = lambda.1se.phi,
#   #                  mse.min   = mse.min.phi,
#   #                  mse.1se   = mse.1se.phi)
#   fold.scan.res = add_row(fold.scan.res,fold = n.fold, type = "phi",
#                    lambda.min = lambda.min.phi,
#                    lambda.1se = lambda.1se.phi,
#                    mse.min   = mse.min.phi,
#                    mse.1se   = mse.1se.phi)
#   
#   proc.time() - ptm
# }


```

## Elastic Net Regression 
### 1.Prepare the Elastic Training Dataset 

